---
title: 
output:
  revealjs::revealjs_presentation:
    self_contained: false
    reveal_plugins: ["search", "zoom", "menu"]
    reveal_options:
    theme: dark
    menu:
      numbers: true

---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_knit$set(echo = FALSE,
                     message = FALSE,
                     warning = FALSE,
                     ggplot2::theme_set(ggplot2::theme_bw()),
                     # root.dir = "F:/Users/aa0856a/ENVS_RS Dropbox/Projects/pix4d/201807XX_Denali_FHM_06/"
                     root.dir = "C:/Users/Minh Nguyen/ENVS_RS Dropbox/Projects/pix4d/201807XX_Denali_FHM_06/"
                     )

library(tidyverse)
library(leaps)
library(ggplot2)
library(cowplot)
library(shiny)
library(shinythemes)
library(png)
library(randomForest)
library(dplyr)
library(stringr)
library(caret)
library(car)
library(splitstackshape)
library(MASS)
library(pracma)
library(philentropy)
library(raster)
library(rgdal)
library(tree)
library(ForestTools)
library(lidR)
library(mapview)
library(corrplot)  #for correlogram
library(e1071)  #for skewness() and kurtosis()
library(gridExtra)  #for grid.arrange()
library(RANN)
library(leaps)
library(jmv) # frequency table/stat
library(glmnet)
```

``` {r echo=FALSE}
raw_df <- read.csv( "../../STV/code/R/AliyaMinh/data/raw_df.csv") %>% 
  mutate(region = as.factor(region),
         status = as.factor(status))

model_df <- raw_df %>%
  group_by(treeid) %>%
  #find color statistics
  mutate(height = max(Z),
         ht_mean = mean(Z),
         ht_skw_Sp = skewness(Z),
         ht_kurt_sp = kurtosis(Z),
         ht_75p = as.numeric(quantile(Z, 0.75)),
         ht_90p = as.numeric(quantile(Z, 0.9)), 
         ht_98p = as.numeric(quantile(Z, 0.98)),
         green_75p = mean(G[Z >= ht_75p]),
         green_90p = mean(G[Z >= ht_90p]),
         green_98p = mean(G[Z >= ht_98p]),
         red_75p = mean(R[Z >= ht_75p]),
         red_90p = mean(R[Z >= ht_90p]),
         red_98p = mean(R[Z >= ht_98p]),
         blue_75p = mean(B[Z >= ht_75p ]),
         blue_90p = mean(B[Z >= ht_90p]),
         blue_98p = mean(B[Z >= ht_98p]),
         greenness_75p = mean(G[Z >= ht_75p]/(R[Z >= ht_75p]+B[Z >= ht_75p])),
         greenness_90p = mean(G[Z >= ht_90p]/(R[Z >= ht_90p]+B[Z >= ht_90p])),
         greenness_98p = mean(G[Z >= ht_98p]/(R[Z >= ht_98p]+B[Z >= ht_98p])),
         redness_75p = mean(R[Z >= ht_75p]/(G[Z >= ht_75p]+B[Z >= ht_75p])),
         redness_90p = mean(R[Z >= ht_90p]/(G[Z >= ht_90p]+B[Z >= ht_90p])),
         redness_98p = mean(R[Z >= ht_98p]/(G[Z >= ht_98p]+B[Z >= ht_98p])),
         blueness_75p = mean(B[Z >= ht_75p]/(G[Z >= ht_75p]+R[Z >= ht_75p])),
         blueness_90p = mean(B[Z >= ht_90p]/(G[Z >= ht_90p]+R[Z >= ht_90p])),
         blueness_98p = mean(B[Z >= ht_98p]/(G[Z >= ht_98p]+R[Z >= ht_98p])),
         med_height = median(Z),
         blueness_mean = mean((B - G)/(B + G)),
         greenness_mean = mean((G - R)/(G + R)),
         redness_mean = mean((R - B)/(R + B)),
         blueness_std = sd((B - G)/(B + G)),
         greenness_std = sd((G - R)/(G + R)),
         redness_std = sd((R - B)/(R + B)),
         blueness_med = median((B - G)/(B + G)),
         greenness_med = median((G - R)/(G + R)),
         redness_med = median((R - B)/(R + B)),
         blueness_skw = skewness((B - G)/(B + G)),
         greenness_skw = skewness((G - R)/(G + R)),
         redness_skw = skewness((R - B)/(R + B)),
         #find overall brightness statistics
         brightness_mean = mean(B + G + R),
         brightness_med = median(B + G + R),
         brightness_std = sd(B + G + R),
         brightness_skw = skewness(B + G + R),
         #find normalized statistics
         red_norm_mean = mean(R / (R + G + B)),
         blue_norm_mean = mean(B / (R + G + B)),
         green_norm_mean = mean(G / (R + G + B)),
         R = mean(R),
         G = mean(G),
         B = mean(B),
         R_ratio = R/(R + G + B ),
         G_ratio = G/(R + G + B ),
         B_ratio = B/(R + G + B ),
         G_R_ratio = G/R,
         G_R_ratio_2 = (G - R)/(G + R))%>% 
  dplyr::select(Id, treeid, status, region, R:G_R_ratio_2)
```


# Analyzing Tree Health Using G-LiHT Data 

Aliya Alimujiang and Minh Nguyen
<br></br>
```{r echo=FALSE, fig.width = 5, fig.height=3}
img <- readPNG("../../STV/code/R/AliyaMinh/images/alaska.png")
grid::grid.raster(img)

```

# Overview
<br></br>

 - Motivation
 - Pre-Processing
    - Descriptive Analytics
    - Feature Extraction
 - Variable Selection
 - Models
 - Conclusion
 - Appendix


## Motivation

<div style="width: 50%; float:left;">

Why tree health?
```{r echo=FALSE, fig.width=6, fig.height=4}
img <- readPNG("../../STV/code/R/AliyaMinh/images/tree_hugger.png")
grid::grid.raster(img)
```

- The future is at stake
- Private and Governments partnership
  - "Green" $2.3T plan
    
</div>

<div style="width: 50%; float:right;">

Why this project?
```{r echo=FALSE, fig.width=6, fig.height=4}
img <- readPNG("../../STV/code/R/AliyaMinh/images/data_algorithms.png")
grid::grid.raster(img)
```

- Great power comes great responsibility - Uncle Ben
- Venture into point cloud

</div>


# Descriptive Analytics

- Read all las files
- Reproject Shp and las files
  - get_las_crowns(las_all, fhm_crowns_shp)

```{r echo=FALSE, fig.width=8, fig.height=5}
img <- readPNG("../../STV/code/R/AliyaMinh/images/preprocessing.png")
grid::grid.raster(img)
```

## Height distribution

```{r echo=FALSE, fig.width=6, fig.height=4}
img <- readPNG("../../STV/code/R/AliyaMinh/images/heightdistribution.png")
grid::grid.raster(img)
```

<table>
  <tr>
    <th> mean</th>
    <th> pct10</th>
    <th> pct30</th>
    <th> median</th>
    <th> pct75</th>
    <th> pct90</th>
    <th> pct98</th>
  </tr>
  <tr>
     <th> 10.33125</th>
    <th> 1.92</th>
    <th> 7.465</th>
    <th> 10.76</th>
    <th> 14.416</th>
    <th> 16.8058</th>
    <th> 21.67</th>
  </tr>
</table>

## Alive vs Dead

```{r echo=FALSE,  fig.width=6, fig.height=3}
img <- readPNG("../../STV/code/R/AliyaMinh/images/alive_dead_pointcloud.png")
grid::grid.raster(img)
```

```{r echo=FALSE,  fig.width=6, fig.height=3}
img <- readPNG("../../STV/code/R/AliyaMinh/images/alivedead.png")
grid::grid.raster(img)
```

## Feature Extraction

How did we go from 6 to 53 variables?

- Structural Variables (X, Y, Z):
  - ht_90p: Height at 90%
  - ht_skw_Sp: Height skewness
- Spectral Variables (R, G, B):
  - greenness_75p: Green ratio at 75% height
  - green_norm_mean: Mean of Green over all Colors
  - G_R_ratio_2: (G - R)/(G + R)

The full list is in the Appendix II.

## Correlation Matrix

```{r echo=FALSE}
img <- readPNG("../../STV/code/R/AliyaMinh/images/corr_plot.png")
grid::grid.raster(img)
```

# Variable selection

<div style="width: 50%; float:left;">
```{r echo=FALSE, fig.width=4, fig.height=6}
img <- readPNG("../../STV/code/R/AliyaMinh/images/descriptive.png")
grid::grid.raster(img)

```
</div>


<div style="width: 50%; float:right;">
<br></br>

- Forward 
- Backward

- n is not much larger than p 
    - a lot of variability
- Bias VS Variance trade off 
</div>

## Forward and Backward
<div style="width: 50%; float:left;">
Forward
```{r echo=FALSE}
img <- readPNG("../../STV/code/R/AliyaMinh/images/forward.png")
grid::grid.raster(img)

```
</div>

<div style="width: 50%; float:right;">
Backward
```{r echo=FALSE}
img <- readPNG("../../STV/code/R/AliyaMinh/images/backward.png")
grid::grid.raster(img)
```
</div>

## Variables Kept

<div style="width: 50%; float:left;">
- Forward: 
   - ht_98p
   - green_98p
    - red_98p
    - brightness_Std
    - red_norm_mean
    - G_ratio
    - G_R_ratio_2
    - green_norm_mean
    - B_ratio
</div>

<div style="width: 50%; float:right;">
- Backward: 
  - red_75p
  - redness_75p
  - redness_98p
  - blue_norm_mean
  - R_ratio
  - G_R_ratio
  - green_norm_mean
  - B_ratio
</div>

# Models

```{r echo=FALSE, fig.width=3, fig.height=2}
img <- readPNG("../../STV/code/R/AliyaMinh/images/predictive_modeling.png")
grid::grid.raster(img)
```

- Parametric
  - Logistic Regression
  - Lasso
- Nonparametric
  - Random Forrest
  - SVM


## Logistic Regression

```{r echo=FALSE,message=FALSE }

# model_df <-model_df %>% group_by(treeid) %>% 
#   filter(row_number()==1) %>%
#   as.data.frame() %>%
#   dplyr::select(-c(treeid, Id, region))
# 
# model_df <- model_df %>%
#   filter(status != "G") %>%
#   mutate(status = factor(status))
# 
# fwd <- model_df%>%
#   dplyr::select(status,ht_98p, green_98p, red_98p, brightness_std, red_norm_mean, G_ratio, G_R_ratio_2, green_norm_mean, B_ratio )
# 
# set.seed(1)
# n= length(fwd$status)
# 
# Z = sample(1:nrow(fwd), 0.7*nrow(fwd))
# fwd_train <- fwd[Z,]
# fwd_test <- fwd[-Z,]
# 
# 
# 
# glm.fits = glm(status~ ., data = fwd_train, family = binomial)
# 
# 
# Prob=fitted.values(glm.fits)
# 
# 
# Probability = predict(glm.fits,fwd_test, type="response")
# 
# Prob=fitted.values(glm.fits)
# 
# # We will classify "Dead trees" as having the probability that exceeds 0.5.
# 
# Predicted.Direction <- ifelse(Probability>0.5,"D","A")
# 
# 
# fwd_conf <- table("Predicted"=Predicted.Direction,"Actual"=fwd_test$status)
# # fwd_conf
```

Fwd Conf Matrix| Actual         
:------------- | :------------- | :-------------
Predicted      | A              | D
A              | 16             | 0
D              | 1              | 7

Accuracy and Kappa

model          | Accuracy       | Kappa
:------------- | :------------- | :-------------
forward        | 0.95833333     | 0.90322581

Accuracy Assessment for fwd_glm model:

Status          | Producer's Accuracy | Omission       | User's Accuracy | Commission
:-------------  | :-------------      | :------------- | :-------------  | :------------- 
Alive           | 0.94117647          | 0.05882353     | 1               | 0
Dead            | 1                   | 0              | 0.875           | 0.125

## Logistic Regression

Accuracy and Kappa

model          | Accuracy       | Kappa
:------------- | :------------- | :-------------
backward       | 0.8750000      | 0.7096774

Accuracy Assessment for bwd_glm model:

Status          | Producer's Accuracy | Omission       | User's Accuracy | Commission
:-------------  | :-------------      | :------------- | :-------------  | :------------- 
Alive           | 0.8823529           | 0.1176471      | 0.9375          | 0.3750 
Dead            | 0.8571429           | 0.1428571      | 0.75            | 0.25


## LASSO
```{r echo=FALSE, fig.width=5, fig.height=3}
img <- readPNG("../../STV/code/R/AliyaMinh/images/lasso_lambda.png")
grid::grid.raster(img)
```

model          | Accuracy       | Kappa
:------------- | :------------- | :-------------
Lasso          | 0.8250000      | 0.61538462

Accuracy Assessment for LASSO  model:

Status          | Producer's Accuracy | Omission       | User's Accuracy | Commission
:-------------  | :-------------      | :------------- | :-------------  | :------------- 
Alive           | 0.95833333          | 0.04166667     | 0.7931034       | 0.4250000
Dead            | 0.625               | 0.375          | 0.90909091      | 0.09090909


## Random Forrest

- Intuition: P is about 2/3 of our observations
- Hyperparameters: mtry, maxnode and ntree
- rf_mtry performed the best originally but rt_fit model had the best result on the test dataset.

<div style="width: 45%; float: left;">
Initial Run

model          | Accuracy       | Kappa
:------------- | :------------- | :-------------
rf_default     | 0.8466667      | 0.6827506
**rf_mtry**    | **0.8766667**  | **0.7424242**
rf_maxnode     | 0.8366667      | 0.6655012
rf_maxtrees    | 0.8233333      | 0.6442890
rf_fit         | 0.8500000      | 0.6878788
</div>

<div style="width: 10%; float: center;">
</div>


<div style="width: 45%; float: right;">
Test Data

model          | Accuracy       | Kappa
:------------- | :------------- | :-------------
rf_default     | 0.9166667      | 0.8139535
rf_mtry        | 0.8750000      | 0.7096774
rf_maxnode     | 0.9166667      | 0.7983193
rf_maxtrees    | 0.9166667      | 0.7983193
**rf_fit**     | **0.9583333**  | **0.9032258**
</div>


## Accuracy Assessment for rf_fit model:

Status          | Producer's Accuracy | Omission       | User's Accuracy | Commission
:-------------  | :-------------      | :------------- | :-------------  | :------------- 
Alive           | 0.94117647          | 0.05882353     | 1               | 0
Dead            | 1                   | 0              | 0.875           | 0.125

Below is the table of the top 5 importance variables for our final model (rf_fit):

Variable       | Importance 
:------------- | :------------- 
G_ratio        | 100.00
greenness_90p  | 64.94
greenness_75p  | 63.44
green_norm_mean| 62.39
G_R_ratio_2    | 54.20

## SVM

- Intuition: Captures complex relationships
- Hyperparameters: Linear/Radial, Cost, and sigma 
- Radial_Grid performed the best originally but the Linear_Grid model had the best result on the test dataset.
<br></br>

model              | Accuracy       | Kappa          | C              | sigma
:-------------     | :------------- | :------------- | :------------- | :-------------
svm_Linear         | 0.7566667      | 0.5128205      | 1.0000000      | NA
svm_Linear_Grid    | 0.8366667      | 0.6724942      | 0.1052632      | NA
svm_Radial         | 0.8500000      | 0.6872960      | 0.5000000      | 0.01763192
**svm_Radial_Grid**| **0.9133333**  | **0.8212121**  | **0.2105263**  | **0.010**

model               | Accuracy       | Kappa
:-------------      | :------------- | :------------- 
svm_Linear          | 0.7916667      | 0.5833333		
**svm_Linear_Grid** | **0.9583333**  | **0.9032258**		
svm_Radial          | 0.9166667      | 0.7983193		
svm_Radial_Grid     | 0.9166667      | 0.7983193

## SVM Accuracy and Kappa

Accuracy Assessment for svm_Linear_Grid model:

Status          | Producer's Accuracy | Omission       | User's Accuracy | Commission
:-------------  | :-------------      | :------------- | :-------------  | :------------- 
Alive           | 0.94117647          | 0.05882353     | 1               | 0
Dead            | 1                   | 0              | 0.875           | 0.125

```{r echo=FALSE, fig.width=6, fig.height=4}
img <- readPNG("../../STV/code/R/AliyaMinh/images/svm_plot.png")
grid::grid.raster(img)
```

# Conclusion

model               | Accuracy       | Kappa
:------------------ | :------------- | :-------------
forward             | **0.95833333** | **0.90322581**
backward            | 0.8750000      | 0.7096774
Lasso               | 0.8250000      | 0.61538462
**rf_fit**          | **0.9583333**  | **0.9032258**
**svm_Linear_Grid** | **0.9583333**  | **0.9032258**	

<br></br>

- Forward performed just as well in **THIS DATASET**
  - Binary question
  - Our recommended model
- Might change with more observations

## Next Steps

- Missing SFM
- Literature Review
  - Feature Extraction
- Gather more data
   - Augmentation
   - Artificial Data


```{r echo=FALSE, fig.width=3, fig.height=3}
img <- readPNG("../../STV/code/R/AliyaMinh/images/next_step.png")
grid::grid.raster(img)
```

## Thank you!

```{r echo=FALSE, fig.width=6, fig.height=4}
img <- readPNG("../../STV/code/R/AliyaMinh/images/yoda.png")
grid::grid.raster(img)
```


# Appendix I – References
 
<small>

- Dr. Alonzo and ENVS-685 Remote Sensing
- NASA G-LiHT data
- Tree species classification from airborne hyperspectral and LiDAR data using 3D convolutional neural networks
- Urban tree species mapping using hyperspectral and lidar data fusion
- Mapping tall shrub biomass in Alaska at landscape scale using structure-from-motion photogrammetry and lidar
- Estimation of standing dead tree class distributions in northwest coastal forests using lidar remote sensing
- Silvi-Net–A dual-CNN approach for combined classification of tree species and standing dead trees from remote sensing data
- Species-related single dead tree detection using multi-temporal ALS data and CIR imagery
- Distinguishing between live and dead standing tree biomass on the North Rim of Grand Canyon National Park
- Detection of Standing Deadwood from Aerial Imagery Products: Two Methods for Addressing the Bare Ground Misclassification Issue
- Individual tree segmentation from a leaf-off photogrammetric point cloud
- Non-destructive aboveground biomass estimation of coniferous trees using terrestrial LiDAR

</small>

# Appendix II – Variables Dictionary

```{r echo=FALSE, fig.width = 10, fig.height=6}
img <- readPNG("../../STV/code/R/AliyaMinh/images/datatable.png")
grid::grid.raster(img)
```

